### 2024-04-29 docs/papers.tex
Set up research workspace.

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tcolorbox}
\usepackage{ifthen}

\tcbuselibrary{skins,breakable}

\title{Journal Article Summaries}
\author{Your Name}
\date{\today}

\colorlet{ArticleBoxBackgroundColorOdd}{white}
\colorlet{ArticleBoxBackgroundColorEven}{gray!10}

\newbool{isEven}
\setbool{isEven}{false}

\newtcolorbox{ArticleColorBox}[1]{%
    breakable, 
    enhanced, 
    colback=\ifbool{isEven}{ArticleBoxBackgroundColorEven}{ArticleBoxBackgroundColorOdd}, 
    colframe=black, 
    fonttitle=\bfseries, 
    title=Article #1
}

\newcounter{ArticleSummaryCounter}
\setcounter{ArticleSummaryCounter}{0}

\newenvironment{ArticleSummary}[1]{
    \refstepcounter{ArticleSummaryCounter} % Increment the counter
    \begin{tcolorbox}[breakable, enhanced, colback=\ifbool{isEven}{ArticleBoxBackgroundColorEven}{ArticleBoxBackgroundColorOdd}, colframe=black, fonttitle=\bfseries, title=Summary \theArticleSummaryCounter: #1]
    \setlength\parskip{0.2cm}
}{
    \end{tcolorbox}
    \vspace*{0.5cm}
    \ifbool{isEven}{\setbool{isEven}{false}}{\setbool{isEven}{true}}
}

\begin{document}

\maketitle

\begin{ArticleSummary}{dfdsf}
\textbf{Citation:} sdfsd

\textbf{Summary:} sdfasf

\textbf{Key Points:}
\begin{itemize}
    \item : sdfs
    \item : 
\end{itemize}

\end{ArticleSummary}
\end{document}